An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-15 paper/paper.tex
Paper: more fixes

\textit{Hashing} is the process of applying a \textit{hash function} to some arbitrary-length message, resulting in a fixed sized \textit{hash}. A good hash function will uniformly distribute the hashes it produces. \textit{Cryptographic hash functions} are hash functions that are suitable for use in cryptography. They possess mathematical properties that allow for confidentiality to be provided, among other properties. To strengthen hashing algorithms, a \textit{salt} can be utilized. A salt is essentially a string of bytes that is randomly picked to be appended to the end of a sequence of data to be hashed. This allows the hashed version of a sequence of data to be different, while still being able to calculate the same hash by applying the same salt. This technique is useful in preventing dictionary attacks, where an attacker has a precomputed list of hashes, and the hash of some data is compared to it in an attempt to find out what the original data is. With a salt, the precomputed hash for the data wll not match, and the list cannot be easily recomputed with the salt, since it is an expensive and lengthy operation.

@@

A \textit{hash table} is a data structure that maps unique keys to values. It has attractive performance features, such as $O(1)$ average lookup time. This means that, on average, the time it takes to retrieve a value from the hash table is independent of the number of entries being stored. This performace comes from the use of a hash function when organizing the data internally. Hash tables also have $O(1)$ insertion and deletion time.

A \textit{distributed hash table} (DHT) is a hash table that is stored across mutliple nodes. Each node is responsible for a portion of the hash table's contents. In exchange for distributing the storage cost among several nodes, DHTs do not have $O(1)$ search, insertion, or deletion time. However, this is generally acceptable because the usefulness of distributing the storage cost far outweights the drawback of slower operations on the data structure. This is especially true since the network latency will likely overshadow the slower performance of the data structure.

@@

DHTs play a major role in many P2P systems. They serve as the means through which peers can find either data or other peers, depending on the system; our system uses a DHT for the latter. In conjunction with the structured overlay network, the DHT allows for peers to discover each other. This is primarily done when looking for peers to send backups to. We discuss how we use our DHT in more detail in section \ref{sec:ourDHT}.

\section{The BitTorrent Protocol} \label{sec:TheBitTorrentProtocol}
BitTorrent is a peer to peer protocol designed for fast file sharing. A peer who wishes to share a file creates a .torrent file containing metadata about the file to be shared. This peer is known as the file's initial \textit{seeder}, and through the use of a BitTorrent client, starts \textit{seeding} the file. When a peer acts as a seeder, its sole purpose is to share the file that it is seeding with other peers. Peers who wish to download the file must acquire its .torrent file through an out-of-band channel, often a torrent indexer. Peers use the .torrent file to locate the initial seeder through either a centralized tracker or a DHT. These peers then connect to the initial seeder and start downloading the file. These peers are known as \textit{leechers} until they possess a full copy of the file, which is when they, themselves, become seeders.

% TODO: add citation and more information about how much better BitTorrent is vs. direct download. Actually, add a lot of citations to this paragraph
While this is more complex than hosting a file on a server and having all users download it, with enough peers there is a huge performance gain with respect to download speeds. The reason this works is due to the differences in upload and download speeds on the Internet; with most Internet connections offered to consumers, the upload bandwidth is much smaller than the download bandwidth. This means that, in the scenario where one machine is tranferring data over the Internet to another machine, download speeds would be severely limited by the upload bandwidth of the uploader. If multiple machines are trying to download from the same uploader, the download speeds will be even worse, as the upload bandwidth is being shared.

% TODO: finish this paragraph
The BitTorrent Protocol takes advantage of this by having leechers download content from multiple peers who are a part of the torrent. A file a partitioned into several file blocks, and a file download is performed by requesting these blocks from multiple peers. % Use combined upload bandwidth, etc.

BitTorrent can transfer files quickly due to multiple seeders each uploading different \textit{chunks} of the file to each leecher. This parallelized download is enhanced by the fact that leechers, too, upload the chunks of the file they possess to other leechers who don't have those chunks. \footnote {This is a simplified version of the BitTorrent protocol. For more information, see \cite{bittorrentProtocol}}

@@

\section{Data Exchange}

>>>>

\textit{Hashing} is the process of applying a \textit{hash function}, which uses a mathematical function, to some arbitrary-length message, resulting in a fixed sized \textit{hash}. Regardless of the input, the hash will usually look like a random string of characters because the mathematical function manipulates the underlying data. This manipulation greatly changes the bits originally used to represent the input data. A good hash function will uniformly distribute the hashes it produces to minimize \textit{collisions}. Collisions are when two inputs hash to the same value, and are often undesired, especially in \textit{cryptographic hash functions}. Cryptographic hash functions are hash functions that are suitable for use in cryptography. They possess mathematical properties that allow for confidentiality to be provided, among other properties.

To strengthen hashing algorithms, a \textit{salt} can be utilized. A salt is essentially a randomly picked string of bytes that is concatenated to data to be hashed. The use of different salts with the same sequence of data creates a different hash (since the total input the hash function is different). The same hash will be calculated if the same salt is used with the same sequence of data. Using salts is a technique useful in preventing \textit{dictionary attacks}, where an attacker compares a precomputed list of inputs and their hashes with a target hash of some unknown data in an attempt to find out what the unknown data is. Dictionary attacks are common on password hashes, because, upon success, the attacker will discover the password used to generate the password hash. With a salt, the precomputed hash for the data wll not match. The list of hashes cannot be easily recomputed with the salt, since hashing is designed to be an expensive and lengthy operation.

@@

A \textit{hash table} is a data structure that maps unique keys to values. It has attractive performance features, such as $O(1)$ average lookup time. This means that, on average, the time it takes to retrieve a value from the hash table is independent of the number of entries being stored. This performace comes from the use of a hash function when organizing the data internally, the details of which is out of the scope of this paper. Hash tables also have $O(1)$ insertion and deletion time.

A \textit{distributed hash table} (DHT) is a hash table that is stored across mutliple nodes. Each node is responsible for a portion of the hash table's contents. In exchange for distributing the storage cost among several nodes, DHTs do not have $O(1)$ search, insertion, or deletion time. However, this is generally acceptable because the usefulness of distributing the storage cost far outweighs the drawback of slower operations on the data structure. This is especially true since the network latency will likely overshadow the slower performance of the data structure.

@@

DHTs play a major role in many P2P systems. They serve as the means through which peers can find either data or other peers, depending on the system; our system uses a DHT for the latter. In conjunction with the structured overlay network, the DHT allows for peers to discover each other. This is primarily done when looking for peers to send backups to. We discuss how we use a DHT in more detail in section \ref{sec:DataExchange}.

\section{The BitTorrent Protocol} \label{sec:TheBitTorrentProtocol}
BitTorrent is a peer to peer protocol designed for fast file sharing. A peer who wishes to share a file creates a .torrent file containing metadata about the file to be shared. This peer is known as the file's initial \textit{seeder}, or uploader, and through the use of a BitTorrent client, starts \textit{seeding}, or uploading, the file. When a peer acts as a seeder, its sole purpose is to share the file that it is seeding with other peers. Peers who wish to download the file must acquire its .torrent file through an out-of-band channel, often a torrent indexer. Peers use the .torrent file to locate the initial seeder through either a centralized tracker or a DHT. These peers then connect to the initial seeder and start downloading the file. These peers are known as \textit{leechers} until they possess a full copy of the file, which is when they, themselves, become seeders.

% TODO: add citation and more information about how much better BitTorrent is vs. direct download. Actually, add a lot of citations to this paragraph
While this is more complex than hosting a file on a server and having users download it, with enough peers there is a huge performance gain with respect to download speeds. The reason this works is due to the differences in upload and download speeds on the Internet; with most Internet connections offered to consumers, the upload bandwidth is much smaller than the download bandwidth. This means that, in the scenario where one machine is tranferring data over the Internet to another machine, download speeds would be severely limited by the upload bandwidth of the uploader. If multiple machines are trying to download from the same uploader, the download speeds will be even worse, as the upload bandwidth is being shared.

% TODO: finish this paragraph
The BitTorrent Protocol takes advantage of this by having leechers download content from multiple peers who are a part of the torrent. A file is partitioned into several \textit{chunks}, or file blocks, and a file download is performed by requesting these chunks from multiple peers. % Use combined upload bandwidth, etc.

BitTorrent can transfer files quickly due to multiple seeders each uploading different chunks of the file to each leecher. This parallelized download is enhanced by the fact that leechers, too, upload the chunks of the file they possess to other leechers who don't have those chunks. \footnote {This is a simplified version of the BitTorrent protocol. For more information, see \cite{bittorrentProtocol}}

@@

\section{Data Exchange} \label{sec:DataExchange}
